Motional Quantum Ground-State Cooing Outside the Lamb-Dicke Regime

### 2017-06-19 paper.tex
Fix conflicts but include edits I made on this branch so far

\begin{document}
\title{Motional Quantum Ground-State Cooing of a Single Sodium Atom}
\author{Yichao Yu}
\author{Nicholas R. Hutzler}
\author{Jessie T. Zhang}
\author{Lee R. Liu}
\author{Kang-Kuen Ni}
\email{ni@chemistry.harvard.edu}
\affiliation{Department of Chemistry and Chemical Biology, Harvard University, Cambridge, Massachusetts, 02138, USA}
\affiliation{Department of Physics, Harvard University, Cambridge, Massachusetts, 02138, USA}
\affiliation{Harvard-MIT Center for Ultracold Atoms, Cambridge, Massachusetts, 02138, USA}

@@

\begin{abstract}
  Single neutral atom trapped in optical tweezer is a promising system for various applications
  in quantum information and simulation.
  The control of the motional states for such a system is crucial
  for exploiting its full potential.
  We report Raman sideband cooling of a single sodium atom to its three-dimensional
  motional ground state in an optical tweezer.
  Despite having a very large Lamb-Dicke parameter, high initial temperature and
  large differential AC Stark shift in the excited state,
  we achieved a ground state preparation fidelity of $77.1(35)\%$ after a few hundred ms of cooling
  by using a few new cooling techniques including addressing high order Raman sideband.
  We demonstrate that Raman sideband cooling to the 3D motional ground state is applicable to
  systems where tight confinement or low initial cooling is hard to achieve.
  The is particularly relevant for systems which are challenging to laser-cool,
  such as molecules and exotic atoms, and opens up a new approach for further cooling
  of these systems.
\end{abstract}

@@

Single neutral atoms trapped in optical tweezers provide a promising system
for various applications including quantum information, quantum chemistry
and quantum simulation of many-body systems.
Long-range interactions between the sites can be achieved by using Rydberg atoms,
polar molecules, or dipolar atoms.
Polar molecules offer the additional advantage of a high degree of tunability,
including the shape, range, and strength of the interaction.
Arrays of optical tweezers offer the ability to move, detect, and manipulate individual sites,
and can be used to prepare complex interacting systems with high fidelity.
In order to achieve long coherence time and full quantum control of the system needed for
these applications,
it is usually necessary to control the thermal motion of the atoms by cooling to the
three dimensional motional ground state in the optical tweezer,
which has already been demonstrated with heavy alkali atoms including rubidium and cesium.
However, some of the important properties of previous experiments,
for example low initial temperature and small Lamb-Dicke parameter,
may not be easily achievable for other systems, such as directly laser-cooled polar molecules,
or atoms with challenging sub-Doppler cooling such as Na, K, and Li.
To extend the versatility of this approach, we must be able to apply it in a more general case.
In this letter, we cool single sodium atoms
trapped in optical tweezers to the motional ground state with Raman sideband cooling.
Despite having a large Lamb-Dicke parameter and high initial temperature, we are able to achieve a ground state probability of $77.1(35)\%$
by utilizing several new cooling techniques and a carefully optimized cooling sequence.

The Raman sideband cooling we used to achieve the high ground state preparation fidelity
consists of multiple cycles of laser pulses to manipulate the internal and
motional states of the atoms.
Figure \ref{f-setup}A shows the schematic of the energy levels and the cooling sequence
in our setup.
Each cooling cycle starts with the sodium atom in the $|F=2, m_F=-2\rangle$
ground electronic state and a certain vibration state $n$.
In the first step, a Raman pulse drives a transition to the motional state $n-\Delta n$
to reduce the motional energy while also changing the internal state to $|F=1, m_F=-1\rangle$.
In the second step, which finishes the cooling cycle,
an optical pumping pulse bring the atom back to the $|F=2, m_F=-2\rangle$ state to take away
the entropy. The second step could also change the motional state of the atoms which
could cause heating. The possibility for this to happen for an atom in the motional level $n$
is approximately proportional to the effective Lamb-Dicke parameter
$\eta^{OP}_{eff}=\sqrt{2n+1}\eta^{OP}$ where $\eta^{OP}=k^{OP}\sqrt{\hbar/2m\omega}$ is the Lamb-Dicke parameter for
optical pumping.
The geometry of the relevant beams and their polarizations is showed in figure \ref{f-setup}B.
The optical tweezer has one weakly confined axial direction (axis $1$) and
two more strongly confined radial directions (axis $2$ and $3$).
Multiple pairs of beams are used to drive the Raman transition during cooling in order to
isolate and maximize the coupling to different trap axis.

Although Raman sideband cooling to motional ground state has already been successfully
implemented on neutral atoms in other experiments using heavier species like Rubidium
\cite{Thompson2013,Kaufman2012}
and Cesium [??], such experiments generally use good polarization gradient cooling
and a tight confinement to achieve a small Lamb-Dicke parameter and effective Lamb-Dicke parameter.
However, this regime is harder to achieve with sodium atoms, creating additional challenges for
us to realize ground state cooling. On the other hand, since these conditions can also be difficult
to meet for other interesting systems like directly laser cooled molecules, the techniques we
use to overcome these challenges could be useful for a wider variety of systems.

>>>>

\begin{document}
\title{Motional Quantum Ground-State Cooing Outside the Lamb-Dicke Regime}
\author{Yichao Yu}
\author{Nicholas R. Hutzler}
\author{Jessie T. Zhang}
\author{Lee R. Liu}
\author{Kang-Kuen Ni}
\email{ni@chemistry.harvard.edu}
\affiliation{Department of Chemistry and Chemical Biology, Harvard University, Cambridge, Massachusetts, 02138, USA}
\affiliation{Department of Physics, Harvard University, Cambridge, Massachusetts, 02138, USA}
\affiliation{Harvard-MIT Center for Ultracold Atoms, Cambridge, Massachusetts, 02138, USA}

@@

\begin{abstract}
  Single neutral atom trapped in optical tweezer is a promising system for various applications
  in quantum information and simulation.
  The control of the motional states for such a system is crucial
  for exploiting its full potential.
  We report Raman sideband cooling of a single sodium atom to its three-dimensional
  motional ground state in an optical tweezer.
  Despite having a very large Lamb-Dicke parameter, high initial temperature and
  large differential AC Stark shift in the excited state,
  we achieved a ground state preparation fidelity of $77(4)\%$ after a few hundred ms of cooling
  by using a few new cooling techniques including addressing high order Raman sideband.
  We demonstrate that Raman sideband cooling to the 3D motional ground state is applicable to
  systems where tight confinement or low initial cooling is hard to achieve.
  The is particularly relevant for systems which are challenging to laser-cool,
  such as molecules and exotic atoms, and opens up a new approach for further cooling
  of these systems.
\end{abstract}

@@

Systems of single neutral atoms trapped in optical tweezers are an exciting platform to study quantum information, quantum chemistry,
and quantum simulation of many-body systems [REF].
This approach provides full single-site resolution and control, and can feature long-range interactions via Rydberg atoms,
polar molecules, or dipolar atoms [REF].
Polar molecules in particular offer the additional advantages of long-lived internal states and a high degree of tunability,
including the shape, range, and strength of the interaction [REF].
Optical tweezers can also be re-arranged in real-time, allowing rapid preparation of atoms in complex geometries with high fidelity [REF].

In order to achieve long coherence times and full quantum control of the system,
it is typically necessary to cool the atom to the
three dimensional motional ground state in the optical tweezer.
This has been demonstrated for neutral Rb and Cs [REF] in experiments that had low initial temperature and small Lamb-Dicke parameter.  
These features
may not be easily achievable for other systems, such as directly laser-cooled polar molecules [REF],
or atoms with challenging sub-Doppler cooling such as Na, K, and Li [REF].
In this letter, we cool single sodium atoms
trapped in optical tweezers to the motional ground state with Raman sideband cooling.
Despite having a large Lamb-Dicke parameter and high initial temperature, we are able to achieve a ground state probability of $77(4)\%$
by utilizing several new cooling techniques and a carefully optimized cooling sequence.  Our approach is quite general, and opens up ground-state cooling for other systems.
